Interpersonal Objectification among Women: Evidence for the Relationship between Being Objectified and Self-Objectification in Same-Gender Interactions

### 2020-11-30 Final_Manuscript.tex
final edits

\documentclass[man]{apa6}
\usepackage{lmodern}
\usepackage{amssymb,amsmath}
\usepackage{ifxetex,ifluatex}
\usepackage{fixltx2e} % provides \textsubscript
\ifnum 0\ifxetex 1\fi\ifluatex 1\fi=0 % if pdftex
  \usepackage[T1]{fontenc}
  \usepackage[utf8]{inputenc}
\else % if luatex or xelatex
  \ifxetex
    \usepackage{mathspec}
  \else
    \usepackage{fontspec}
  \fi
  \defaultfontfeatures{Ligatures=TeX,Scale=MatchLowercase}
\fi
% use upquote if available, for straight quotes in verbatim environments
\IfFileExists{upquote.sty}{\usepackage{upquote}}{}
% use microtype if available
\IfFileExists{microtype.sty}{%
\usepackage{microtype}
\UseMicrotypeSet[protrusion]{basicmath} % disable protrusion for tt fonts
}{}
\usepackage{hyperref}
\hypersetup{unicode=true,
            pdftitle={Objectification in Action: Self- and Other-Objectification in Same-gender Interactions},
            pdfauthor={Randi L. Garcia, Kat Kyuchukova, Asha Hinson, \& Laneé Jung},
            pdfkeywords={Interpersonal sexual objectification, self-objectifiation, authenticity,
actual interactions, psychology of women, dyadic data analysis},
            pdfborder={0 0 0},
            breaklinks=true}
\urlstyle{same}  % don't use monospace font for urls
\usepackage{graphicx,grffile}
\makeatletter
\def\maxwidth{\ifdim\Gin@nat@width>\linewidth\linewidth\else\Gin@nat@width\fi}
\def\maxheight{\ifdim\Gin@nat@height>\textheight\textheight\else\Gin@nat@height\fi}
\makeatother
% Scale images if necessary, so that they will not overflow the page
% margins by default, and it is still possible to overwrite the defaults
% using explicit options in \includegraphics[width, height, ...]{}
\setkeys{Gin}{width=\maxwidth,height=\maxheight,keepaspectratio}
\IfFileExists{parskip.sty}{%
\usepackage{parskip}
}{% else
\setlength{\parindent}{0pt}
\setlength{\parskip}{6pt plus 2pt minus 1pt}
}
\setlength{\emergencystretch}{3em}  % prevent overfull lines
\providecommand{\tightlist}{%
  \setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}
\setcounter{secnumdepth}{0}
% Redefines (sub)paragraphs to behave more like sections
\ifx\paragraph\undefined\else
\let\oldparagraph\paragraph
\renewcommand{\paragraph}[1]{\oldparagraph{#1}\mbox{}}
\fi
\ifx\subparagraph\undefined\else
\let\oldsubparagraph\subparagraph
\renewcommand{\subparagraph}[1]{\oldsubparagraph{#1}\mbox{}}
\fi

@@

  \title{Objectification in Action: Self- and Other-Objectification in
Same-gender Interactions}
    \author{Randi L. Garcia\textsuperscript{1, 2}, Kat
Kyuchukova\textsuperscript{1, 2}, Asha Hinson\textsuperscript{1}, \&
Laneé Jung\textsuperscript{2}}
    \date{}

\shorttitle{Objectification in Action}
\affiliation{
\vspace{0.5cm}
\textsuperscript{1} Department of Psychology, Smith College\\\textsuperscript{2} Program in Statistical and Data Sciences, Smith College}
\keywords{Interpersonal sexual objectification, self-objectifiation, authenticity, actual interactions, psychology of women, dyadic data analysis}
\usepackage{csquotes}
\usepackage{upgreek}
\captionsetup{font=singlespacing,justification=justified}

@@

\authornote{Randi L. Garcia is in the Psychology Department
and the Statitsical and Data Sciences program at Smith College.

@@

\abstract{
Existing research on interpersonal objectification has focused mostly on
links between objectification, self-objectification, and negative
outcomes for women within mixed-gender interactions (Garcia, Earnshaw,
\& Quinn, 2016; Gervais, Sáez, Riemer, \& Klein, 2020). The purpose of
the present study was to extend past research on interpersonal
objectification to interactions between women. Women were brought into
the laboratory and interacted face-to-face in same-sex dyads. Dyadic
analysis (Kenny, Kashy, \& Cook, 2006) was utilized to detect whether
partners' objectification of each other affected state
self-objectification, as well as the resulting feelings of comfort and
authenticity during the interaction. Results revealed a significant
positive relationship between being objectified by a woman interaction
partner and women's own self-objectification (a partner effect). There
was no significant relationship between self-objectification and
interaction inauthenticity, but there were significant negative effects
of inauthenticity on career aspirations and relationship agency (actor
effects). The significant partner effect of objectification on
self-objectification suggests that women being objectified by other
women may result in feelings of self-objectification, and this finding
opens up questions for future research about how interpersonal
objectification might function in situations without an immediate male
gaze.

@@

Fredrickson and Roberts (1997)'s Objectification Theory, suggests that,
in addition to being steeped in a culture that objectifies women, women
are objectified in actual interpersonal encounters. The negative effects
of this interpersonal objectification for women have been theorized to
be the strongest under the \enquote{male gaze,} that is, when it is a
perceived or actual man doing the objectifying (Calogero, 2004; Gay \&
Castano, 2010; Gervais, Vescio, \& Allen, 2011), but what are the
consequences when a woman objectifies another woman in an interaction?
Psychological researchers studying the sexual objectification of women
have in recent years explored the interpersonal process of the
objectification, finding evidence for \emph{self}-objectification in the
target as a proximal consequence of being objectified (Garcia et al.,
2016; Riemer, Sáez, Brock, \& Gervais, 2020; Strelan \& Pagoudis, 2018).
In a recent review, Gervais et al. (2020) organized the extant
literature on interpersonal self-objectification and proposed a
theoretical model called the Social Interaction Model of Objectification
(SIMO). The SIMO focuses on understanding mixed-gender interactions,
acknowledging the patriarchal power structure embedded in these mixed
interactions, but can it be extended to interactions among women?
Although there is evidence that women can objectify other women
(Gervais, Holland, \& Dodd, 2013; Loughnan et al., 2015; Puvia \& Vaes,
2013), studies investigating the process of interpersonal
objectification in interactions between women is scarce.

The current study uses a face-to-face interaction methodology to being
to answer a series of research questions about interpersonal
objectification among women: Is objectification by a female interaction
partner related to an increase in self-objectification for the woman
being objectified? Does objectification by a woman have the same
downstream negative consequences for women as being objectified by a
male interaction partner? The current study addresses these questions by
investigating interpersonal objectification among interacting pairs
where both partners identify as women.

@@

Self-objectification is a psychological process that translates the
experiences of interpersonal objectification (Gervais et al., 2020;
Loughnan, Baldissarri, Spaccatini, \& Elder, 2017) into negative mental
health outcomes (e.g., interaction inauthenticity, anxiety, lower
self-esteem, and poor cognitive performance; Calogero, Tantleff-Dunn, \&
Thompson, 2011; Moradi \& Huang, 2008). Calogero et al. (2011) proposes
that self-objectification can be conceptualized as a learned trait, or
\emph{trait} self-objectification (TSO), whereby one adopts a habitual
third-person perspective on one's own appearance. Furthermore,
self-objectification can also be elicited momentarily, for example, when
viewing sexualized images in movies and magazines (Morry \& Staska,
2001), when trying on sexualizing clothing (B. L. Fredrickson, Roberts,
Noll, Quinn, \& Twenge, 1998), or during video chat when you believe
your interaction partner can only see your body (not face; Saguy, Quinn,
F Dovidio, \& Pratto, 2010). This momentary self-objectification is
referred to as \emph{state} self-objectification (SSO; Calogero et al.,
2011; Moradi \& Huang, 2008) and is characterized by feeling like a body
rather than a full self within a particular moment, instance, or
context.

@@

Past research investigating women objectifying other women has focused
on the psychological conditions that lead women to objectify other
women. Harsey and Zurbriggen (2020) found that (trait)
self-objectification was related to the objectification of women to a
similar degree for male and female participants. Along the same lines,
Strelan and Hargreaves (2005) found that the more women self-objectify,
the more they objectify other women. Could the process work in the other
direction as well, that is, could objectification by a woman produce
(state) self-objectification, in a similar manner to objectification by
a man? Hill and Fischer (2008) assessed women's experiences of
objectification from men independently from women's experiences of
objectification from other women. They found that women may be
socialized not only to see themselves as objects, but perhaps to see
other women as objects as well. This process, however, was not confirmed
to occur during an actual interpersonal interaction. Thus, there is
evidence that women do objectify other women, indeed, Loughnan et al.
(2017) found that women objectify other women to a greater extent than
they objectify men. Women have also been found to objectify
(dehumanized) sexualized targets presented as images (Puvia \& Vaes,
2013), but very little is known about interpersonal sexual
objectification among women and what the effects of this objectification
might be. Past studies have found that objectification has more adverse
consequences for women than men (Gervais et al., 2011; Moradi \& Huang,
2008; Saguy et al., 2010), however, we do not know much about the effect
of the gender of the \emph{objectifyer} on these potential detrimental
outcomes.

@@

Studies have shown that within social encounters women are gazed at more
than men (Henley, 1977), often times feel \enquote{looked at} within
interpersonal interactions (Argyle \& Williams, 1969), and are likely to
internalize the objectifying gaze on their physical self (Puvia \& Vaes,
2013). As mentioned above, interpersonal objectification has been found
to have negative consequences for women. Gervais et al. (2011) found
that an objectifying gaze by a male interaction partner (confederate)
was associated with lower math performance for women than being
objectified by a female interaction partner. Gervais et al. (2020)
recently reviewed the research on interpersonal objectification and
organized our current theoretical understanding of this process in the
SIMO. What is clear from Gervais et al. (2020)'s review is that we know
little about women-on-women interpersonal objectification.

Past research has found that the experience of state
self-objectification in mixed-gender contexts (stranger and romantic;
Strelan and Pagoudis (2018); Meltzer (2020); Zurbriggen, Ramsey, and
Jaworski (2011)) has negative consequences for women, but what about in
the context of interactions with other women? Is the
self-objectification experienced in mixed-gender interactions associated
with the same negative process as that experienced in same-gender
interactions? The ample research demonstrating that the male gaze has a
particularly detrimental effect for women would suggest no (Calogero,
2004; Gay \& Castano, 2010; Gervais et al., 2011; Saguy et al., 2010),
but, there is evidence that women do objectify other women (Harsey \&
Zurbriggen, 2020; Loughnan et al., 2015; Puvia \& Vaes, 2013) and recent
experimental evidence suggests that the female gaze causes no less
self-objectification than the male gaze (Yilmaz \& Bozo, 2019). It would
be helpful to understand the consequences of this \emph{intra}group
objectification for women. When women objectify other women, does it
lead to self-objectification in the same way men's objectification of
women does (Garcia et al., 2016)? If so, does the self-objectification
experienced in these same-gender interactions have the same negative
consequences for authenticity in that interaction?

There is some evidence that reduced authenticity is a consequence of
self-objectification (Garcia et al., 2016; Terán, Jiao, \& Aubrey,
2020). A recent study has investigated the effect of
self-objectification on the reduction of relationship building skills in
general (including same-sex friendships) (Terán et al., 2020).
Authenticity reduction and interaction quality disruptions have also
been found in research on stigmatized-stigmatizer interactions (M. R.
Hebl \& Dovidio, 2005; Pearson et al., 2008; Richeson \& Shelton, 2003;
Shelton, Richeson, \& Salvatore, 2005), and we might view the experience
of being objectified in an interaction as a potential identity threat
situation (Gervais et al., 2020; Nadal \& Haynes, 2012). Further,
empirical evidence reveals that healthy relationship functioning
manifests through authenticity in romantic relationships (Brunell et
al., 2010), an intergroup context at least in heterosexual
relationships. Evidence for other adverse consequences of interpersonal
objectification include reductions in career aspirations (Garcia et al.,
2016) and a decrease in concentration and impairment in female cognitive
performance (Kahalon, Shnabel, \& Becker, 2018; D. M. Quinn, Chaudoir,
\& Kallen, 2011). These are also negative consequences found for women
under stereotype threat (Davies, Spencer, \& Steele, 2005). When a woman
is objectified by a man, and subsequently experiences
self-objectification, the intergroup nature and power differential of
this encounter might trigger threat. Thus, perhaps there are fewer
negative consequences when a woman is objectified by another woman. That
is, the intergroup threat literature might predict that woman-woman
interpersonal objectification processes diverge from mixed-gender
interpersonal objectification processes precisely because they are not
intergroup interactions (at least with respect to gender identity).

In summary, a woman objectified by another woman may not be having the
same negative consequences that cascade from situations that trigger
group-based identity threat (Deaux \& Major, 1987; Dovidio, Hebl,
Richeson, \& Shelton, 2006; M. R. Hebl \& Dovidio, 2005); however,
feeling like a body rather than a full human (higher SSO) in any
interaction, whether intergroup or intragroup, may be enough to reduce
women's feelings of authenticity and social competence, regardless of
the gender of the objectifyer (Terán et al., 2020; Tolman, Impett,
Tracy, \& Michael, 2006).

@@

In the current study, we sought to examine what occurs during an
interaction in which one or both partners are objectifying each other,
similarly to Garcia et al. (2016), but between same-gender interpersonal
interactions among women. Moreover, the current study uses a
face-to-face interaction paradigm and dyadic data analysis techniques to
examine the relationships for both women simultaneously. Although the
literature on intragroup objectification among women is small, the
results are mixed, and they do not cover interpersonal encounters, we
expected to replicate some of the results found in Garcia et al. (2016).
Most importantly, we predicted that being objectified by one's
interaction partner would be related to state self-objectification
(SSO). We also expected that TSO would moderate this relationship,
amplifying the positive association between being objectified and SSO
for women higher in TSO. Here we would like to note that Puvia and Vaes
(2013) would alternatively predict that women's tendency to
self-objectify (TSO) leads them to objectify other women (more
precisely, to dehumanize a sexualized woman). In addition, Strelan and
Hargreaves (2005) and Harsey and Zurbriggen (2020) would also predict
the TSO to other-objectification link, further theorizing that this
relationship is mediated by the woman's own state self-objectification
(SSO). Thus, the limited but extant literature on woman-woman
objectification points to a possible alternative model from the model
tested in the current study. Where appropriate, we report results
considering the causal directions implied by this alternative mediation
model.

@@

\emph{All} participants were told that they were assigned to the
\enquote{College Relationships} condition. Self-objectification has been
found to occur after a mere relationship prime among women (Sanchez \&
Broccoli, 2008) because, in Western culture and beyond, women need to
look attractive to obtain and maintain successful relationships, thus,
the \enquote{College Relationships} condition may heighten
self-objectification and the evaluation of other women in a sexualized
way. The decision was made to ask even heterosexual women to judge their
fellow-woman partners for their potential as romantic partners. We felt
that this prompt would keep the study closest to a replication of the
previous Garcia et al. (2016) version of the study, and strengthen the
potential for objectification that would normally be low in the context
of a psychology laboratory. Recall that, past research has found that
heterosexual women are indeed able to evaluate other women's potential
as romantic partners (i.e., their sexual attractiveness; Puvia \& Vaes,
2013)---women may be unfortunately quite used to thinking about their
own \enquote{datability,} and we suspect this this habitual thought
pattern will translate to their thoughts about other women. Analyses
comparing levels of other-objectification between heterosexual and
non-heterosexual women are provided in the measures section below.

@@

\begin{table}[tbp]
\begin{center}
\begin{threeparttable}
\caption{\label{tab:corrtable}Correlations and Descriptive Statistics among Study Variables}
\begin{tabular}{llllllllll}
\toprule
 & \multicolumn{1}{c}{$M$} & \multicolumn{1}{c}{$SD$} & \multicolumn{1}{c}{1.} & \multicolumn{1}{c}{2.} & \multicolumn{1}{c}{3.} & \multicolumn{1}{c}{4.} & \multicolumn{1}{c}{5.} & \multicolumn{1}{c}{6.} & \multicolumn{1}{c}{7.}\\
\midrule
1. Actor's trait self objectification (TSO) & -0.35 & 2.64 &  &  &  &  &  &  & \\
2. Actor's objectification of partner & -1.58 & 1.21 & .20 &  &  &  &  &  & \\
3. Partner's objectification of the actor & -1.58 & 1.21 & .09 & .14 &  &  &  &  & \\
4. Actor's state self-objectification & 1.92 & 1.13 & .13 & -.09 & .30* &  &  &  & \\
5. Actor's authenticity of interaction & 5.23 & 1.02 & -.02 & -.07 & .03 & -.10 &  &  & \\
6. Actor's future relationship agency & 4.69 & 0.96 & .04 & .09 & -.10 & -.09 & .23+ &  & \\
7. Actor's cognitive performance & 5.03 & 2.29 & .08 & .11 & .06 & .02 & .11 & .07 & \\
8. Actor's career aspirations & 3.88 & 0.61 & .05 & -.14 & -.08 & .01 & .33** & .26* & -.07\\
\bottomrule
\end{tabular}
\end{threeparttable}
\end{center}
\end{table}

>>>>

\documentclass[man]{apa6}
\usepackage{lmodern}
\usepackage{amssymb,amsmath}
\usepackage{ifxetex,ifluatex}
\usepackage{fixltx2e} % provides \textsubscript
\ifnum 0\ifxetex 1\fi\ifluatex 1\fi=0 % if pdftex
  \usepackage[T1]{fontenc}
  \usepackage[utf8]{inputenc}
\else % if luatex or xelatex
  \ifxetex
    \usepackage{mathspec}
  \else
    \usepackage{fontspec}
  \fi
  \defaultfontfeatures{Ligatures=TeX,Scale=MatchLowercase}
\fi
% use upquote if available, for straight quotes in verbatim environments
\IfFileExists{upquote.sty}{\usepackage{upquote}}{}
% use microtype if available
\IfFileExists{microtype.sty}{%
\usepackage{microtype}
\UseMicrotypeSet[protrusion]{basicmath} % disable protrusion for tt fonts
}{}
\usepackage{hyperref}
\hypersetup{unicode=true,
            pdftitle={Interpersonal Objectification among Women: Evidence for the Relationship between Being Objectified and Self-Objectification in Same-Gender Interactions},
            pdfauthor={Randi L. Garcia, Kat Kyuchukova, Asha Hinson, \& Laneé Jung},
            pdfkeywords={Interpersonal sexual objectification, self-objectifiation, authenticity,
actual interactions, psychology of women, dyadic data analysis},
            pdfborder={0 0 0},
            breaklinks=true}
\urlstyle{same}  % don't use monospace font for urls
\usepackage{graphicx,grffile}
\makeatletter
\def\maxwidth{\ifdim\Gin@nat@width>\linewidth\linewidth\else\Gin@nat@width\fi}
\def\maxheight{\ifdim\Gin@nat@height>\textheight\textheight\else\Gin@nat@height\fi}
\makeatother
% Scale images if necessary, so that they will not overflow the page
% margins by default, and it is still possible to overwrite the defaults
% using explicit options in \includegraphics[width, height, ...]{}
\setkeys{Gin}{width=\maxwidth,height=\maxheight,keepaspectratio}
\IfFileExists{parskip.sty}{%
\usepackage{parskip}
}{% else
\setlength{\parindent}{0pt}
\setlength{\parskip}{6pt plus 2pt minus 1pt}
}
\setlength{\emergencystretch}{3em}  % prevent overfull lines
\providecommand{\tightlist}{%
  \setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}
\setcounter{secnumdepth}{0}
% Redefines (sub)paragraphs to behave more like sections
\ifx\paragraph\undefined\else
\let\oldparagraph\paragraph
\renewcommand{\paragraph}[1]{\oldparagraph{#1}\mbox{}}
\fi
\ifx\subparagraph\undefined\else
\let\oldsubparagraph\subparagraph
\renewcommand{\subparagraph}[1]{\oldsubparagraph{#1}\mbox{}}
\fi

@@

  \title{Interpersonal Objectification among Women: Evidence for the Relationship
between Being Objectified and Self-Objectification in Same-Gender
Interactions}
    \author{Randi L. Garcia\textsuperscript{1, 2}, Kat
Kyuchukova\textsuperscript{1, 2}, Asha Hinson\textsuperscript{1}, \&
Laneé Jung\textsuperscript{2}}
    \date{}

\shorttitle{Objectification among Women}
\affiliation{
\vspace{0.5cm}
\textsuperscript{1} Department of Psychology, Smith College\\\textsuperscript{2} Program in Statistical and Data Sciences, Smith College}
\keywords{Interpersonal sexual objectification, self-objectifiation, authenticity, actual interactions, psychology of women, dyadic data analysis}
\usepackage{csquotes}
\usepackage{upgreek}
\captionsetup{font=singlespacing,justification=justified}

@@

\authornote{Randi L. Garcia is in the Psychology Department
and the Program in Statistical and Data Sciences at Smith College.

@@

\abstract{
Existing research on interpersonal objectification has focused mostly on
links between sexual objectification, self-objectification, and negative
outcomes for women within mixed-gender interactions (Garcia, Earnshaw,
\& Quinn, 2016; Gervais, Sáez, Riemer, \& Klein, 2020). The purpose of
the present study was to extend Garcia et al. (2016)`s past research on
mixed-gender interpersonal objectification to interactions between
women. Women were brought into the laboratory and interacted
face-to-face in same-sex dyads. Dyadic analysis (Kenny, Kashy, \& Cook,
2006) was utilized to detect whether partners' objectification of each
other affected state self-objectification, as well as the resulting
feelings of comfort and authenticity during the interaction. Results
revealed a significant positive relationship between being objectified
by a woman interaction partner and a woman's own self-objectification (a
partner effect). There was no significant relationship between
self-objectification and interaction inauthenticity, but there were
significant negative effects of inauthenticity on career aspirations and
relationship agency (actor effects). The significant partner effect of
objectification on self-objectification suggests that women being
objectified by other women may result in feelings of
self-objectification, and this finding opens up questions for future
research about how interpersonal objectification might function in
situations without an immediate male gaze.

@@

Fredrickson and Roberts (1997)'s Objectification Theory suggests that,
in addition to being steeped in a culture that objectifies women, women
are objectified in actual interpersonal encounters. The negative effects
of this interpersonal objectification for women have been theorized to
be the strongest under the \enquote{male gaze,} that is, when it is a
perceived or actual man doing the objectifying (Calogero, 2004; Gay \&
Castano, 2010; Gervais, Vescio, \& Allen, 2011). What are the
consequences when a woman objectifies another woman in an interaction?
Psychological researchers studying the sexual objectification of women
have in recent years explored the interpersonal process of the
objectification, finding evidence for \emph{self}-objectification in the
target as a proximal consequence of being objectified (Garcia et al.,
2016; Riemer, Sáez, Brock, \& Gervais, 2020; Strelan \& Pagoudis, 2018).
However, this research has largely focused on mixed-gender interactions
only. In a recent review, Gervais et al. (2020) organized the extant
literature on interpersonal self-objectification and proposed a
theoretical model called the Social Interaction Model of Objectification
(SIMO). The SIMO focuses on understanding mixed-gender interactions,
acknowledging the patriarchal power structure embedded in these mixed
interactions, but can this model be extended to interactions among
women? Although there is evidence that women can objectify other women
(Gervais, Holland, \& Dodd, 2013; Loughnan et al., 2015; Puvia \& Vaes,
2013), studies investigating the process of interpersonal
objectification in interactions between women is scarce.

The current study uses the face-to-face interaction methodology deployed
in Garcia et al. (2016) to begin to answer a series of research
questions about interpersonal objectification among women: Is
objectification by a female interaction partner related to an increase
in self-objectification for the woman being objectified? Does
objectification by a woman have the same downstream negative
consequences for women as being objectified by a male interaction
partner? The current study addresses these questions by investigating
interpersonal objectification among interacting pairs where both
partners identify as women.

@@

Self-objectification is a psychological process that translates the
experiences of interpersonal objectification (Gervais et al., 2020;
Loughnan, Baldissarri, Spaccatini, \& Elder, 2017) into negative mental
health outcomes (e.g., interaction inauthenticity, anxiety, lower
self-esteem, and poor cognitive performance; Calogero, Tantleff-Dunn, \&
Thompson, 2011; Moradi \& Huang, 2008). Calogero et al. (2011) proposes
that self-objectification can be conceptualized as a learned trait, or
\emph{trait} self-objectification (TSO), whereby one adopts a habitual
third-person perspective on one's own appearance. Furthermore,
self-objectification can also be elicited momentarily, for example, when
viewing sexualized images in movies and magazines (Morry \& Staska,
2001), when trying on sexualizing clothing (B. L. Fredrickson, Roberts,
Noll, Quinn, \& Twenge, 1998), or during video chat when you believe
your interaction partner can only see your body and not your face
(Saguy, Quinn, Dovidio, \& Pratto, 2010). This momentary
self-objectification is referred to as \emph{state} self-objectification
(SSO; Calogero et al., 2011; Moradi \& Huang, 2008) and is characterized
by feeling like a body rather than a full self within a particular
moment, instance, or context.

@@

Past research investigating women objectifying other women has focused
on the psychological conditions that lead women to objectify other
women. Harsey and Zurbriggen (2020) found that (trait)
self-objectification was related to the objectification of women to a
similar degree for male and female participants. Along the same lines,
Strelan and Hargreaves (2005) found that the more women self-objectify,
the more they objectify other women. Could the process work in the other
direction as well? That is, could objectification by a woman produce
(state) self-objectification, in a similar manner to objectification by
a man? Hill and Fischer (2008) assessed women's experiences of
objectification from men independently from women's experiences of
objectification from other women. They found that women may be
socialized not only to see themselves as objects, but perhaps to see
other women as objects as well. This process, however, was not confirmed
to occur during an actual interpersonal interaction. Thus, there is
evidence that women do objectify other women, indeed, Loughnan et al.
(2017) found that women objectify other women to a greater extent than
they objectify men. Women have also been found to objectify
(dehumanized) sexualized targets presented as images (Puvia \& Vaes,
2013), but very little is known about interpersonal sexual
objectification among women and what the effects of this objectification
might be. Past studies have found that objectification has more adverse
consequences for women than men (Gervais et al., 2011; Moradi \& Huang,
2008; Saguy et al., 2010), however, we do not know much about the effect
of the gender of the \emph{objectifier} on these potential detrimental
outcomes.

@@

Studies have shown that within social encounters women are gazed at more
than men (Henley, 1977), often times feel \enquote{looked at} within
interpersonal interactions (Argyle \& Williams, 1969), and are likely to
internalize the objectifying gaze on their physical self (Puvia \& Vaes,
2013). As mentioned above, interpersonal objectification has been found
to have negative consequences for women. Gervais et al. (2011) found
that an objectifying gaze by a male interaction partner (confederate)
was associated with lower math performance for women than being
objectified by a female interaction partner. Gervais et al. (2020)
recently reviewed the research on interpersonal objectification and
organized our current theoretical understanding of this process in the
SIMO. What is clear from Gervais et al. (2020)'s review is that we know
little about women-on-women interpersonal objectification, in
unstructured contexts.

Past research has found that the experience of state
self-objectification in mixed-gender contexts (Meltzer, 2020; stranger
and romantic; Strelan \& Pagoudis, 2018; Zurbriggen, Ramsey, \&
Jaworski, 2011) has negative consequences for women, but what about in
the context of interactions with other women? Is the
self-objectification experienced in mixed-gender interactions associated
with the same negative process as that experienced in same-gender
interactions? The ample research demonstrating that the male gaze has a
particularly detrimental effect for women would suggest no (Calogero,
2004; Gay \& Castano, 2010; Gervais et al., 2011; Saguy et al., 2010).
However, there is evidence that women do objectify other women (Harsey
\& Zurbriggen, 2020; Loughnan et al., 2015; Puvia \& Vaes, 2013) and
recent experimental evidence suggests that the female gaze causes no
less self-objectification than the male gaze (Yilmaz \& Bozo, 2019). It
would be helpful to understand the consequences of this
\emph{intra}group objectification for women. When women objectify other
women, does it lead to self-objectification in the same way men's
objectification of women does (Garcia et al., 2016)?

If so, does the self-objectification experienced in these same-gender
interactions have the same negative consequences for authenticity in
that interaction?

One mechanism by which interpersonal objectification translates into
negative consequences is interaction inauthenticity - when actors feel
that they are not being their true selves in an interaction. There is
some evidence that reduced authenticity is a consequence of
self-objectification (Garcia et al., 2016; Terán, Jiao, \& Aubrey,
2020). A recent study has investigated the effect of
self-objectification on the reduction of relationship building skills in
general, including same-sex friendships (Terán et al., 2020).
Authenticity reduction and interaction quality disruptions have also
been found in research on stigmatized-stigmatizer interactions (M. R.
Hebl \& Dovidio, 2005; Pearson et al., 2008; Richeson \& Shelton, 2003;
Shelton, Richeson, \& Salvatore, 2005), and we might view the experience
of being objectified in an interaction as a potential identity threat
situation (Gervais et al., 2020; Nadal \& Haynes, 2012). Further,
empirical evidence reveals that healthy relationship functioning
manifests through authenticity in romantic relationships (Brunell et
al., 2010), at least in an intergroup context such as heterosexual
romantic relationships. Evidence for other adverse consequences of
interpersonal objectification include reductions in career aspirations
(Garcia et al., 2016) and a decrease in concentration and impairment in
female cognitive performance (Kahalon, Shnabel, \& Becker, 2018; D. M.
Quinn, Chaudoir, \& Kallen, 2011). These are also negative consequences
found for women under stereotype threat (Davies, Spencer, \& Steele,
2005; Spencer, Steele, \& Quinn, 1999). When a woman is objectified by a
man, and subsequently experiences self-objectification, the intergroup
nature and power differential of this encounter might trigger threat.
Thus, me might expect there to fewer negative consequences when a woman
is objectified by another woman. That is, the intergroup threat
literature might predict that woman-woman interpersonal objectification
processes diverge from mixed-gender interpersonal objectification
processes precisely because they are not intergroup interactions, at
least with respect to gender identity.

In summary, a woman objectified by another woman may not be having the
same negative consequences that cascade from situations that trigger
group-based identity threat (Deaux \& Major, 1987; Dovidio, Hebl,
Richeson, \& Shelton, 2006; M. R. Hebl \& Dovidio, 2005); however,
feeling like a body rather than a full human (higher SSO) in any
interaction, whether intergroup or intragroup, may be enough to reduce
women's feelings of authenticity and social competence, regardless of
the gender of the objectifier (Terán et al., 2020; Tolman, Impett,
Tracy, \& Michael, 2006).

@@

In the current study, we sought to examine what occurs during an
interaction in which one or both partners are potentially objectifying
each other and where both interaction partners identify as women.
Moreover, the current study used the same face-to-face interaction
paradigm as Garcia et al. (2016) and dyadic data analysis techniques to
examine the relationships for both women simultaneously. Although the
literature on intragroup objectification among women is small, the
results are mixed, and they do not cover actual interpersonal
encounters, we expected to replicate some of the results found in Garcia
et al. (2016). Most importantly, we predicted that being objectified by
one's interaction partner would be related to state self-objectification
(SSO). We also expected that TSO would moderate this relationship,
amplifying the positive association between being objectified and SSO
for women higher in TSO. Here we would like to note that Puvia and Vaes
(2013) would alternatively predict that women's tendency to
self-objectify (TSO) leads them to objectify other women (more
precisely, to dehumanize a sexualized woman). In addition, Strelan and
Hargreaves (2005) and Harsey and Zurbriggen (2020) would also predict
the TSO to other-objectification link, further theorizing that this
relationship is mediated by the woman's own state self-objectification
(SSO). Thus, the limited but extant literature on woman-woman
objectification points to a possible alternative model from the model
tested in the current study. Where appropriate, we report results
considering the causal directions implied by this alternative mediation
model.

@@

\emph{All} participants were told that they were assigned to the
\enquote{College Relationships} condition. Self-objectification has been
found to occur after a mere relationship prime among women (Sanchez \&
Broccoli, 2008) because, in Western culture and beyond, women need to
look attractive to obtain and maintain successful relationships, thus,
the \enquote{College Relationships} condition may heighten
self-objectification and the evaluation of other women in a sexualized
way. The decision was made to ask all women participating in the study,
even heterosexual women, to judge their fellow-woman partners for their
potential as romantic partners. We felt that this prompt would keep the
study closest to a replication (and extension) of the previous Garcia et
al. (2016) version of the study, and strengthen the potential for
objectification that would normally be low in the context of a
psychology laboratory. Recall that, past research has found that
heterosexual women are indeed able to evaluate other women's potential
as romantic partners, specifically, their sexual attractiveness (Puvia
\& Vaes, 2013). Women may be unfortunately quite used to thinking about
their own \enquote{datability,} and we suspected that this habitual
thought pattern would translate to their thoughts about other women.
Analyses comparing levels of other-objectification between heterosexual
and non-heterosexual women are provided in the measures section below.

@@

\begin{table}[tbp]
\begin{center}
\begin{threeparttable}
\caption{\label{tab:corrtable}Correlations and Descriptive Statistics among Study Variables}
\begin{tabular}{llllllllll}
\toprule
 & \multicolumn{1}{c}{$M$} & \multicolumn{1}{c}{$SD$} & \multicolumn{1}{c}{1.} & \multicolumn{1}{c}{2.} & \multicolumn{1}{c}{3.} & \multicolumn{1}{c}{4.} & \multicolumn{1}{c}{5.} & \multicolumn{1}{c}{6.} & \multicolumn{1}{c}{7.}\\
\midrule
1. Actor's trait self objectification (TSO) & -0.35 & 2.64 &  &  &  &  &  &  & \\
2. Actor's objectification of partner & -1.58 & 1.21 & .20 &  &  &  &  &  & \\
3. Partner's objectification of the actor & -1.58 & 1.21 & .09 & .14 &  &  &  &  & \\
4. Actor's state self-objectification & 1.92 & 1.13 & .13 & -.09 & .30* &  &  &  & \\
5. Actor's authenticity of interaction & 5.23 & 1.02 & -.02 & -.07 & .03 & -.10 &  &  & \\
6. Actor's future relationship agency & 4.69 & 0.96 & .04 & .09 & -.10 & -.09 & .23+ &  & \\
7. Actor's cognitive performance & 5.03 & 2.29 & .08 & .11 & .06 & .02 & .11 & .07 & \\
8. Actor's career aspirations & 3.88 & 0.61 & .05 & -.14 & -.08 & .01 & .33** & .26* & -.07\\
\bottomrule
\addlinespace
\end{tabular}
\begin{tablenotes}[para]
\normalsize{\textit{Note.} * p < .05. ** p < .01.}
\end{tablenotes}
\end{threeparttable}
\end{center}
\end{table}

\begin{lltable}
